### 2019-03-20 10d_SYM_Notes.tex
Added comments on Baulieu's scalar supersymmetry

\documentclass[10pt, oneside]{article}

\input ./math_headers.sty

\newcommand{\map}{\ul{\mr{Map}}}
\newcommand{\dBRST}{\mathrm d_{\mathrm{BRST}}}
\renewcommand{\d}{\mathrm{d}}
\newcommand{\sD}{\slashed{D}}
\newcommand{\sdel}{\slashed{\partial}}
\newcommand{\cO}{\mathcal{O}}
\newcommand{\sdelbar}{\overline{\slashed{\partial}}}
\newcommand{\st}[1]{{}^*{#1}}

\newcommand{\brian}[1]{\textcolor{blue}{BW: #1}}

\addbibresource{Twist.bib}

\title{Notes on 10d Super Yang-Mills}
\author{Chris Elliott, Pavel Safronov and Brian Williams}

\date{\today}

\begin{document}

\maketitle

\section{Minimal Super Yang-Mills}
Our starting point will be the theory complexifying the usual 10d super Yang-Mills theory.  Fix a complex reductive gauge group $G$ with Lie algebra $\gg$.  The ordinary fields of super Yang-Mills theory on $\RR^{10}$ consist of a boson: a connection $A$ on the trivial $G$-bundle, and a fermion: a $\gg$-valued section $\lambda$ of the Weyl spinor bundle associated to the spinor representation $S_+$ \footnote{If we didn't complexify we would instead consider $G_\RR$ a compact connected Lie group, and a section of the Majorana-Weyl spinor bundle, which necessitates working in Lorentzian signature.  I think for our purposes it's interesting enough to just consider the complexified theory and avoid signature issues.  The complexified theory twists to holomorphic Chern-Simons theory with complex gauge group.}.  These fields are acted upon by the group of gauge transformations -- $G$-valued functions on $\RR^{10}$.

We can model the stack of fields modulo gauge transformations infinitesimally near the point $0$ by the corresponding BRST complex.  This is the local super Lie algebra
\[L_{\mr{BRST}} = \Omega^0(\RR^{10}; \gg) \to \Omega^1(\RR^{10}; \gg) \oplus \Omega^0(\RR^{10}; \Pi S_+ \otimes \gg)\]
with the de Rham differential, placed in cohomological degrees 0 and 1, with bracket induced from the Lie bracket on $\gg$.

The action functional in 10d super Yang-Mills is given by
\[S(A,\lambda) = \int_{\RR^{10}} \langle F_A \wedge \ast F_A + (\lambda, \sd D_A \lambda)\rangle,\]
where $\langle - \rangle$ denotes an invariant pairing on $\gg$, and $(,)$ denotes a scalar-valued pairing $S_+ \otimes S_- \to \CC$ (there will be a unique such pairing, up to rescaling, characterized by the condition that $(\rho(v)\lambda_1,\rho(v)\lambda_2) = (\lambda_1,\lambda_2)$ for each $v \in \CC^{10}$, where $\rho$ denotes Clifford multiplication).

We can re-encode this data in terms of the classical BV complex (Phil and I wrote this down in \cite[Section 3.1]{ElliottYoo1}).  This is the $L_\infty$-algebra whose underlying cochain complex takes the form
\[\xymatrix{
\Omega^0(\RR^{10}; \gg) \ar[r]^{\d} &\Omega^1(\RR^{10}; \gg) \ar[r]^{\d \ast \d} &\Omega^9(\RR^{10}; \gg) \ar[r]^{\d} &\Omega^{10}(\RR^{10}; \gg) \\
&\Omega^0(\RR^{10}; \Pi S_- \otimes \gg) \ar[r]^{\ast \sd \d} &\Omega^{10}(\RR^{10}; \Pi S_- \otimes \gg), &
}\]
with degree $-3$ invariant pairing induced by the invariant pairing on $\gg$ and the pairing $(,)$ between $S_+$ and $S_-$, and with degree 2 and 3 brackets given by the action of $\Omega^0(\RR^{10}; \gg)$ on everything along with
\begin{align*}
\ell_2^{\mr{Bos}} \colon \Omega^1(\RR^{10};\gg) \otimes \Omega^1(\RR^{10};\gg) &\to \Omega^{9}(\RR^{10};\gg) \\
(A \otimes B) &\mapsto [A \wedge \ast \mr d B] + [\ast \mr d  A \wedge B] + \mathrm{d} \ast[A \wedge B] \\
\ell_2^{\mr{Fer}} \colon \Omega^1(\RR^{10};\gg) \otimes \Omega^0(\RR^{10}; S_{+} \otimes \gg) &\to \Omega^{10}(\RR^{10}; S_{-} \otimes \gg) \\
(A \otimes \lambda) &\mapsto \ast \sd A \lambda
\end{align*}
in degree 2, and the map
\begin{align*}
\ell_3 \colon \Omega^1(\RR^{10};\gg) \otimes \Omega^1(\RR^{10};\gg) \otimes \Omega^1(\RR^{10};\gg) &\to \Omega^{9}(\RR^{10};\gg) \\
(A \otimes B \otimes C) &\mapsto [A \wedge \ast[B \wedge C]] + [B \wedge \ast[C \wedge A]] + [C \wedge \ast[A \wedge B]]
\end{align*}
in degree 3.

\subsection{On-Shell Supersymmetry Action}
We can define an action of the 10d $\mc N=1$ supersymmetry algebra on the complex of BRST fields in this minimal super Yang-Mills theory.  The Poincar\'e action is clear, and the action of the supersymmetry $Q$ is generated -- in the usual Physics notation -- by the transformation \cite{BrinkSchwarzScherk}
\begin{align*}
\delta_Q A &= \Gamma(Q,\lambda) \\
\delta_Q \lambda &= \sd F_A Q,
\end{align*}
where the notation $\sd F_A$ stands for the iterated Clifford multiplication $\sd F_A = F_{ij} \gamma^i \gamma^j$.  To check that this defines a supersymmetry action we need to check it's compatible with the brackets in the supersymmetry algebra.  So, we calculate
\begin{align*}
[\delta_{Q_1}, \delta_{Q_2}] A &= (\Gamma(Q_2,\sd F_A Q_1) - \Gamma(Q_1,\sd F_A Q_2)) \\
&=  F_{ij}(Q_2 \gamma^k \gamma^i \gamma^j Q_1 - Q_1 \gamma^k \gamma^i \gamma^j Q_2)\\
&=  F_{ij}(Q_2 \gamma^k \gamma^i \gamma^j Q_1 - Q_2 \gamma^j \gamma^i \gamma^k Q_1)\\
&=  F_{ij}(Q_2 \gamma^k \gamma^j \gamma^i Q_1 - Q_2 \gamma^j \gamma^k \gamma^i Q_1)\\
&=  F_{ij}\delta^{kj}(Q_2 \gamma^i Q_1) \\
&= \delta_{[Q_1, Q_2]} A,
\end{align*}
where on the third line we used the fact that the pairing $\Gamma(-,-)$ is symmetric -- i.e. that $\lambda_1 \gamma^i \lambda_2 = \lambda_2 \gamma^i \lambda_1$ -- three times, and on the fourth and fifth lines we used the Clifford relations.  Similarly we can calculate
\begin{align*}
[\delta_{Q_1}, \delta_{Q_2}] \lambda &= (\sd F_{\Gamma(Q_2, \lambda)} Q_1 - \sd F_{\Gamma(Q_1,\lambda)} Q_2) \\
&= (Q_2 \gamma_j \dd_i \lambda + [Q_2\gamma_i \lambda, Q_2 \gamma_j \lambda]) (\gamma^i \gamma^j Q_1) - (1 \leftrightarrow 2) \\
&= 
\end{align*}

\begin{remark}
I'm part way through trying to do these calculations myself, but for reference I found these calculations discussed in the master's thesis \cite{Guillen} we've discussed before.  For instance for the commutator of two supersymmetries acting on a spinor see equation 2.1.13.
\end{remark}

%Failure to commute with the classical BV differential

\subsection{An Off-Shell Supersymmetry Action?}
\brian{I realized I'm still confused with the role of Baulieu's $h$-field. 
I really suspect it's different than the usual auxiliary field story. 
It seems like Baulieu doesn't even write down the apparent off-shell SUSY module structure. 
}
\brian{I think without ever introducing an auxiliary field we can obtain the following module structure.}

\def\oloc{\mathcal{O}_{loc}}

Let $\oloc^{BRST}$ and $\oloc^{BV}$ be the local functionals for the BRST and BV fields, respectively. 
The BRST operator endows $\oloc^{BRST}$ with the structure of a cochain complex, and the BV operator together with the BV bracket endow $\oloc^{BV}[-1]$ with the structure of a dg Lie algebra. 
Note that there is a map of dg Lie algebras
\[
\oloc^{BV}[-1] \to {\rm End}(\oloc^{BRST})
\]
sending a functional $I$ to the endomorphism $\{I,-\}$.
\brian{No there is not.}

\def\cN{\mathcal{N}}

The cursory definition of the linear map
\[
\mf g_{\cN=1} \to {\rm End}(\oloc^{BRST})
\]
is not a map of Lie algebras. 
It fails to preserve the Lie bracket by a term proportional to the equations of motion in the field $\lambda$. 
I claim that Baulieu's introduction of the $h$-field does not resolve this issue. (We can see this by counting fermion number. 
The space where the failure for this to be a Lie map is $S_+$, yet we are only adding a scalar $h$ in the auxiliary, so there's no way.

Instead, what one should try is the following. 
(This doesn't ever use a holomorphic language, so maybe I'm making a silly mistake.)
There is an obvious lift of the linear map $\mf g_{\cN=1} \to {\rm End}(\oloc^{BRST})$ to the BV complex
\[
\mf g_{\cN=1} \to \oloc^{BV} [-1] .
\]
%The prescription for defining this map is simple. 
This is still, of course, not a map of dg Lie algebras.
However, I claim that there is an $L_\infty$ correction to this map.

Fix a basis $\{Q_\alpha\}$ of $S_+$ and write a general element of the form $Q_{\alpha} = \epsilon^\alpha Q_\alpha$.
The putative, linear action, of the element $Q$ on the BV complex is through functionals of the form
\[
\int \epsilon \lambda^* \sd F_A 
\]
where $\lambda^*$ denotes the anti-field to $\lambda$. 
I claim that we can correct the action by adding a quadratic term to the action that sends a pair $Q_1 \otimes Q_2$ to the functional
\[
\int \epsilon_1 \lambda^* \epsilon_2 \lambda^*  .
\]

In terms of an $L_\infty$ action, I'm saying that there is the usual linear map
\[
I^{(1)} : \mf g_{\cN=1} \to \oloc^{BV} 
\]
plus an $L_\infty$ correction of the form:
\[
I^{(2)} : \mf g_{\cN=1}^{\otimes 2} \to \oloc^{BV} [-2]  .
\]
You may ask why we have to stop there, but I don't have a great conceptual answer besides the computational fact that it seems like it works out. o
I claim that $I^{(1)} + I^{(2)}$ determines an off-shell action of $\mf g_{\cN=1}$ on the BV complex.

At the level of generators, we are proposing a new transformation law of the form 
\begin{align*}
\delta A &= \d c + \Gamma(\epsilon,\lambda) \\
\delta \lambda &= [c,\lambda] + \epsilon \sd F_A  \pm \# \epsilon (\epsilon \lambda^*)  \\
\delta \lambda^* &= \sd \partial \lambda + [c, \lambda^*] + \epsilon \sd A^*  .
\end{align*}
I've included terms coming from the BV operator (which have no $\epsilon$ dependence). 
The only new term is the last one in the transformation law for $\lambda$, $\epsilon (\epsilon \lambda^*)$.

\chris{
I'm a little confused by a few things: can I ask for clarification?
\begin{enumerate}
 \item Firstly, while I understand the map $(I \mapsto \{I, -\}) \colon \oloc^{BV}[-1] \to {\rm End}(\oloc^{BV})$, I don't understand how it lifts to ${\rm End}(\oloc^{BRST})$.  For instance (maybe a little heuristically, but I hope what I'm trying to say makes sense), let $\alpha$ be a scalar field and let $\alpha^*$ be its antifield.  Let $I = (\alpha^*)^2$ in $\oloc^{BV}$.  Then $\{(\alpha^*)^2, \alpha\} = 2\alpha^*$ is a non-trivial element of $\oloc^{BV}$ which is not in the image of $\oloc^{BRST}$, which says that my na\"ive guess for how to define your map doesn't work, so I must be supposed to do something more clever?
 \item I wanted to clarify something about your ``counting fermion number'' argument.  I thought that this argument only applied to the action of a whole supersymmetry algebra, not a single square zero supercharge.  That is, if we're trying to construct off-shell BRST supersymmetry then the space of BRST fields should yield a representation of the supersymmetry algebra.  Representations $V_0 \oplus V_1$ of the supersymmetry algebra have to have the same number of bosonic and fermionic degrees of freedom -- i.e. $\dim V_0 = \dim V_1$, because we can choose a supertranslation $Q$ which squares to a translation.  So if $\alpha$ denotes the supersymmetry algebra action, the composite of $\alpha(Q)|_{V_0} \colon V_0 \to V_1$ and $\alpha(Q)|_{V_1} \colon V_1 \to V_0$ is equal to the action of a translation, which is an isomorphism, and therefore $\alpha(Q)$ must also be an isomorphism, so $\dim V_0 = \dim V_1$.  However this argument doesn't apply anymore if you just want an action of a single square-zero odd symmetry.  Maybe you have something else in mind though?
 \item Could you include some of the calculation that the map you define at the end is an $L_\infty$ map?  This seems like it might be exactly what we're looking for.
\end{enumerate}
}

\brian{
\begin{enumerate}
\item I was being sloppy, and you are right. 
Think of BRST as functions on $B \mf g$ and BV as functions on $T^* B\mf g$.
Since vector fields include inside poly-vector fields there is a map of dg Lie algebras
\[
{\rm Der}(\cO(B \mf g)) = {\rm Vect}(B \mf g) \to \cO(T^* B \mf g) .
\]
So, really, the map should go the other way.
I'll fix the above when I have more time, but basically what I'm saying is that even though we can't write down a Lie map to derivations of BRST, we can map one to the bigger BV complex.
That's why the action looks like its by poly-vector fields, not just derivations. 
\item
I think I was also being sloppy here, and I don't see a way of making my argument precise. 
I will try to find the reference (I think it may be Berkovits) where he says something to this effect. 
\item
OK, I wanted to see if you thought it was reasonable before including the details. 
I'll work on writing that up now!
\end{enumerate}
}

\subsubsection{Abelian case}

The BV differential is
\begin{align*}
\delta_{BV} c &= 0 \\
\delta_{BV} A &= \d c \\
\delta_{BV} \lambda &= 0 \\
\delta_{BV} \lambda^* &= \sd D_A \lambda \\
\delta_{BV} A^* &= \d\ast F \\
\delta_{BV} c^* &= \d A^*
\end{align*}

\[I^{(1)} = \int (\lambda^*, \sd F \epsilon) + \int A^* \Gamma(\epsilon, \lambda)\]

\[\delta_{BV} I^{(1)} = \int (\sd D_A\lambda, \sd F\epsilon) + \int (\d\ast F) \Gamma(\epsilon, \lambda).\]
\section{Baulieu's 10 Model}

Baulieu \cite{Baulieu} considers an extension of 10d Super-Yang Mills including an auxiliary $\gg$-valued scalar field $h$.  The inclusion of this field breaks the manifest $\SO(10)$ symmetry to the subgroup $\SU(5)$, corresponding to a choice of complex structure on $\RR^{10}$.  Let's describe this classical theory in an explicitly $\SU(5)$-invariant way.  With respect to this choice of complex structure $\CC^5$ becomes Calabi-Yau: we'll denote the holomorphic top-form by $\Omega$ and the map $\Omega^{p,q}(\CC^5) \to \Omega^{p-1,q-1}(\CC^5)$ induced from the K\"ahler structure by $J$.

The ordinary fields $A$ and $\lambda$ of super Yang-Mills will decompose according to the decomposition of the vector and Weyl spinor representations of $\SO(10)$ into irreducible $\SU(5)$-representations.  So, explicitly $A$ splits up into fields
\[A_{1,0} + A_{0,1} \in \Omega^{1,0}(\CC^5; \gg) \oplus \Omega^{0,1}(\CC^5; \gg),\]
and $\lambda$ splits up into fields
\[\chi + \psi + B \in \Pi(\Omega^0(\CC^5; \gg) \oplus \Omega^{1,0}(\CC^5; \gg) \oplus \Omega^{0,2}(\CC^5; \gg)).\]
In terms of these fields, and including the auxiliary field $h$, the action functional becomes
\[S(A_{1,0}, A_{0,1}, \chi, \psi, B, h) = \int \langle J^2(F_{2, 0}\wedge F_{0,2})\Omega - \frac 12 \|h\|^2\Omega + h J(F_{1,1}) \Omega - J( \chi \wedge (\ol \dd_{A_{0,1}} \psi)) \Omega - J^2(B \wedge (\dd_{A_{1,0}} \psi) - (B \wedge \ol \dd_{A_{0,1}} B)) \Omega \rangle.\]
One derives this action functional by decomposing the 10d super Yang-Mills action functional into $\SU(5)$ irreducible component fields, then introducing a Lagrange multiplier $h$ to eliminate the $F_{1,1}^2$ term.
%SU(5) BRST and BV Fields, Action Functional, BV Complex

\subsection{Off-Shell Action of a Scalar Supersymmetry}

Consider the following square-zero supersymmetry acting on Baulieu's theory:
\begin{align*}
\delta A_{1,0} &= \psi \\
\delta \chi &= h \\
\delta B &= F_{0,2} \\
\end{align*}
acting trivially on all other (BRST) fields.  This clearly squares to zero, so defines an action of the supergroup $\Pi \CC$ on the BRST complex.  We can calculate the variation of the action functional with respect to this symmetry, and see that it vanishes.  Indeed,
\begin{align*}
\delta S &= \int \delta \langle J^2(F_{2, 0}\wedge F_{0,2})\Omega - \frac 12 \|h\|^2\Omega + h J(F_{1,1}) \Omega - J( \chi \wedge (\ol \dd_{A_{0,1}} \psi)) \Omega - J^2(B \wedge (\dd_{A_{1,0}} \psi) - (B \wedge \ol \dd_{A_{0,1}} B)) \Omega \rangle \\
&= \int \langle J^2(\dd_{A_{1,0}} B \wedge F_{0,2})\Omega - 0 + h J(\ol \dd_{A_{0,1}} \psi) \Omega - J( h (\ol \dd_{A_{0,1}} \psi)) \Omega - J^2(F_{0,2} \wedge (\dd_{A_{1,0}} \psi) - 0) \Omega \rangle \\
&= 0.
\end{align*}
This should mean that the $\Pi \CC$ action is off-shell, so extends to an action on the full BV complex (the above calculation should be equivalent to showing that $\delta$ commutes with the BV differential).

\section{Homotopy Data}
%We were going to check in particular that everything was a chain map

\subsection{Homotopy Transfer of the Scalar Supersymmetry}

\pagestyle{bib}
\printbibliography

\end{document}