### 2019-12-10 matter.tex
matter

\documentclass[10pt, oneside]{article}

\input ./combined_macros.sty

\newcommand{\Cl}{\mathrm{Cl}}
\newcommand{\Dens}{\mathrm{Dens}}

\addbibresource{Twist.bib}

\usepackage{pdflscape}
\usetikzlibrary{shapes.geometric, arrows, positioning}

\tikzstyle{s16} = [rectangle, rounded corners, minimum width=1.8cm, minimum height=1cm,text centered, draw=black,fill=red!30]
\tikzstyle{s16chiral} = [s16, dashed]
\tikzstyle{s8 } = [rectangle, rounded corners, minimum width=1.8cm, minimum height=1cm,text centered, draw=black,fill=orange!30]
\tikzstyle{s4} = [rectangle, rounded corners, minimum width=1.8cm, minimum height=1cm,text centered, draw=black,fill=yellow!30]
\tikzstyle{s2chiral} = [rectangle, dashed, rounded corners, minimum width=1.8cm, minimum height=1cm,text centered, draw=black,fill=green!30]
\tikzstyle{dimension} = [circle, text centered, text width=0.7cm, minimum height=0.7cm, draw=black]
\tikzstyle{arrow} = [thick,->,>=stealth]

\title{Matter, uniformly}
\author{Chris Elliott\and Pavel Safronov \and Brian Williams}

\date{\today}

\begin{document}

\maketitle

\begin{definition}
Let $\fg$ be a Lie algebra.
An even (resp. odd) {\bf supersymmetric matter pair} is a pair $(\SS, U)$ where 
\begin{itemize}
\item[(1)] $\SS$ is a $\ZZ/2$-graded spinorial representation equipped with an even (resp. odd) nondegenerate $V$-valued pairing
\[
\Gamma : \wedge^2(\SS)^0 \xto{\cong} V
\]
(resp. $\Gamma : \wedge^2(\SS)^1 \xto{\cong} V$), and

\item[(2)] $U$ is an even (resp. odd) symplectic $\fg$-representation.
\end{itemize}
\end{definition}

\begin{lem}
For a fixed Lie algebra $\fg$, a complete classification of the supersymmetric matter pairs \brian{Should we say minimal here? What about dimension 3 and 2?} (for dimensions $n \geq 2$) are:
\begin{itemize}
\item[(1)] Dimension $n = 4$. 
There is an odd supersymmetric matter pair $(\SS = S_+ \oplus S_-, U = R \oplus R^*)$ where the odd pairing $\Gamma : S_+ \otimes S_- \xto{\cong} V$ is the usual $\Gamma$ pairing. Here, $R$ is a $\fg$-representation, and in the odd symplectic representation $U = R \oplus R^*$ we view $R$ as even and $R^*$ as odd. 
\item[(2)] Dimension $n=6$.
There are two even supersymmetric matter pairs: $(\SS = S_+ \otimes W_+, U)$ and $(\SS = S_- \otimes W_-, U)$ where $W_\pm$ is a $2$-dimension symplectic vector space and $U$ is a symplectic $\fg$-representation. 
The even nondegenerate pairings are the usual $\Gamma$ pairings $\Gamma : \wedge^2 (S_\pm) \xto{\cong} V$ 
\end{itemize}
\end{lem}

\begin{prop}
\brian{how to state $3 \psi$-rule}

\begin{itemize}
\item[(1)] Dimension $n = 4$.
Suppose $Q_+ \oplus Q_- \in S_+ \oplus S_-$.
Then, the following diagram commutes
\[
\xymatrix{
S_+ \otimes S_- \ar[dr]_-{(Q_+ + Q_-,-)} \ar[rr]^-{\Gamma} & & V \ar[dl]^-{\rho(v)} \\
& S_+ \oplus S_- &  
}
\]
\item[(2)] Dimension $n = 6$. 
Suppose $Q \in S_\mp$.
Then, the following diagram commutes
\[
\xymatrix{
\wedge^2 S_\pm \ar[dr]_-{(Q,-)} \ar[rr]^-{\Gamma} & & V \ar[dl]^-{\rho(v)} \\
& S_\pm & 
}
\]
\end{itemize}
\end{prop}

To a supersymmetric matter pair we will associate a free BV theory whose underlying BRST fields consist of:
\begin{itemize}
\item a scalar $\phi \in C^\infty(\RR^n ; \SS^*)$;
\item a negative Weyl spinor valued in $U$; $\psi \in C^\infty(\RR^4 ; S_- \otimes U)$ .
\begin{dfn}
The BRST theory associated to a supersymmetric matter pair (even or odd) $(\SS, U)$ has underlying bundle of BRST fields:
\[

\end{document}