### 2020-02-24 old stuff/anomaly.tex

\documentclass[10pt, oneside]{article}

\input ./combined_macros.sty

\addbibresource{Twist.bib}

\title{Twists of supersymmetric gauge theories}
\author{Chris Elliott\and Pavel Safronov \and Brian Williams}

\date{\today}

\begin{document}

\maketitle

\section{One-loop anomalies}

\subsection{Holomorphic Chern-Simons}

\begin{prop}
Let $X$ be any flat Calabi-Yau manifold of odd complex dimension $d$. 
The one-loop anomaly to quantizing holomorphic Chern-Simons theory with values in $\fg$ on $X$ is proportional to the local cocycle
\[
\int_X {\rm Tr}_{\rm ad} (\alpha \underbrace{\partial \alpha \cdots \partial \alpha}_{d}) \in \cloc^\bu \left(\sE_{\rm hCS}\right) .
\] 
where the trace is taken in the adjoint representation of $\fg$. 
\end{prop}

\begin{rmk}
If we expand $\alpha$ in components as $\alpha = c + A^{0,1} + \sum_{i \geq 2} A^{0,i}$ where $c \in C^\infty(X) \otimes \fg$ and $A^{0,i} \in \Omega^{0,i}(X) \otimes \fg$, then up to some nonzero factor we can write the anomaly cocycle as 
\[
\int_X {\rm Tr}_{\rm ad} (c \partial A^{0,1} \cdots \partial A^{0,1}) = \int {\rm Tr}_{\rm ad} \left(c (F_{A^{0,1}})^d \right) .
\]
\end{rmk}

\end{document}