### 2022-04-15 twist_selecta_referee_comments.tex
A fistful of corrections

\documentclass[10pt, oneside]{article}

\input ./combined_macros.sty

\addbibresource{Twist.bib}

\title{A Taxonomy of Twists of Supersymmetric Yang--Mills Theory -- Comments on Referee Report}
\author{Chris Elliott\and Pavel Safronov \and Brian Williams}

\date{\today}

\begin{document}
\maketitle

We are very grateful to the referee for their numerous helpful comments.  Please find below a summary of the changes we have made in response to the report.

\begin{enumerate}
 \item Page 8: Added an explanatory comment.
 \item Page 18: \chris{I propose that the best thing to do here would be to add a longish remark on $\infty$-categories with some references, but reassuring the reader that no knowledge of details is needed to understand the content of the paper.}
 \pavel{Hmm. I would prefer not to add anything about $\infty$-categories as we are not really using them. Could be good to say that we use the equivalence as a black box: we \emph{define} FMPs as $L_\infty$ algebras. Like in Definition 1.6, how about we just give a definition ``an FMP is an $L_\infty$ algebra'' and add a remark after the definition that it is justified by the Lurie--Pridham result, but we will not be using it?}

 \item Page 20 points 1 and 2: Some further explanation was added for the notation here. \pavel{We define cotangent bundles etc of formal moduli problems, but then work with local formal moduli problems. The confusion arises because we never really defined what $T^*BL$ is when $L$ is a local $L_\infty$ algebra.}
 \item Page 20 point 3: Added a remark, thank you.
 \item Page 27: Added an additional remark here to clarify the definition.
 \brian{Maybe the issue is also with us calling this a BRST 'theory'? Perhaps less conflicting will be to refer to this as 'BRST data'. What do you think?} \chris{I don't think that's the issue, but maybe we should use a letter other than $\mathcal M$ for the formal moduli problem here.  Like maybe $\mathcal F$ for fields?}
 \pavel{Perhaps, we should just reply to the referee rather than add a remark? Agree about changing the letters though to clarify matters: the FMP he writes down is associated to BV fields, while here we have an extra data. If we change the letters, it will be easier to address the objection.}
 \item Page 29: Used $1/j$ for the root of the canonical bundle instead of $1/m$.
 \item Page 32: Clarified the condition $p^*S_N = S_M$ with a remark and fixed a small typo in the definition.
 \item Page 34 point 1: The intended map here was the pullback along $\mr{Re} \colon V \to V_\RR$, corrected here.
 \item Page 34 point 2: The parameter $\eps$ in the given equivalence corresponds to $\d z$, where $z$ is a complex coordinate on the factor $\CC$.  Because this parameter has degree 1, it is formal and the issue indicated here does not arise.
 \item Page 35: Added some clarification to the notation here.
 \item Page 36: Yes, $Q_i$ and $\lambda$ are sections of the same thing, clarified this point. \chris{rest to respond to}
 \item Page 37, point 1: We have added the clarifying adjective ``Euclidean''.
 \item Page 37, point 2: We have clarified when we are working with merely alternative algebras and when we require associativity.
 \item Page 38: We have mentioned that it is a version of the Fierz identity, thank you.
 \item Page 51: Thank you, reference added.
 \item Page 52: Thank you for the remark; we have removed the confusing sentence.
 \item Page 54: Added a reference for this claim to the relevant section of [ES].
 \item Page 56: This section was written confusingly, and we have corrected it.  Both $Q_+$ and $Q_-$ individually have stabilizers isomorphic to $\mr{SU}(4)$, and the intersection of these stabilizers is isomorphic to $\mr{SU}(3)$.
 \item Page 77: Thanks, we have corrected the confusing appearance of ``six-dimensional'' which refers to the spin group of the 6-dimensional vector space, rather than the dimension of the spin representation which is 4.
\end{enumerate}

\end{document}